---
title: " "
output:
  bookdown::word_document2:
    reference_docx: results_formatting.docx
linkcolor: black
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
library(bookdown)
Sys.setenv(FISHBASE_API="sealifebase")
#knitr::opts_chunk$set(echo = TRUE)
```

```{r Temp_BVC, include = FALSE}
source("tempBVC.R")
```

### Results {.unnumbered}

Richness during summer for all stations was generally stable, while winter was more variable in the southern estuaries than in the northern estuaries (Figure \@ref(fig:NPlot)).

```{r NPlot, echo=FALSE, message=FALSE, fig.width=7.5, fig.height=5, fig.cap = "An amazing plot"}

plot(NPlot)

```

However, total abundance over time was quite variable for all systems in the winter season, with summer generally rather stable.

```{r AbundPlot, echo=FALSE, message=FALSE, fig.width=7.5, fig.height=5, fig.cap = "Another amazing plot"}

plot(AbundPlot)

```

Seagrass coverage did stuff (Figure \@ref(fig:BVCPlot)).

```{r BVCPlot, echo=FALSE, message=FALSE, fig.width=7.5, fig.height=5, fig.cap = "Yet another amazing plot"}

plot(BVCPlot)

```

Temperature was interesting.

```{r TempPlot, echo=FALSE, message=FALSE, fig.width=7.5, fig.height=5, fig.cap = "One more amazing plot"}

plot(TempPlot)

```

```{r Models, include = FALSE}
source("RichnessModel.R")
```

Is there any association between changes in abundance and changes in temperature and/or seagrass coverage? GLMMs used to determine. Formula used for all models is: `abund ~ year_Z + temp_Z +  bvc_Z + offset(log(n_hauls)) + ar1(yearMonth + 0|systemZone)` where abund is ... z is ...

Here's what the abundance model output looks like for each system and season combination in Table \@ref(tab:abundModels)

```{r abundModels, echo=FALSE}

emphasize.strong.cells(which(abundPvalues < 0.05, arr.ind = TRUE))
pander(abundEstimates,
             caption = "(\\#tab:abundModels) The first table.")

# knitr::kable(test,
#              caption = "The first table.")

```

And this is for richness models as in Table \@ref(tab:richModels)

```{r richModels, echo=FALSE}

emphasize.strong.cells(which(richPvalues < 0.05, arr.ind = TRUE))
pander(richEstimates,
             caption = "(\\#tab:richModels) The second table.")

# knitr::kable(test,
#              caption = "The first table.")

```

Looking closer at temp, specifically as in [@Miner2021], the upper and lower 10th percentiles of temp are sometimes changing. Mostly in winter, though not significant using linear model alone (Figure \@ref(fig:upperLower)).

```{r upperLower, echo=FALSE, message=FALSE, fig.width=7.5, fig.height=5, fig.cap = "Still another amazing plot"}

plot(annualTemp)

```

This is the abundance results when run thru AIC (Table \@ref(tab:aicAbund)).

```{r aicAbund, echo=FALSE}

emphasize.strong.cells(which(tempAPvalues < 0.05, arr.ind = TRUE))
pander(tempAEstimates,
             caption = "(\\#tab:aicAbund) The second table.")

# knitr::kable(test,
#              caption = "The first table.")

```

This is the richness results when run thru AIC (Table \@ref(tab:aicRich)).

```{r aicRich, echo=FALSE}

emphasize.strong.cells(which(tempNPvalues < 0.05, arr.ind = TRUE))
pander(tempNEstimates,
             caption = "(\\#tab:aicRich) The second table.")

# knitr::kable(test,
#              caption = "The first table.")

```

Now for some multivariate exploration to determine which taxa are underlying these environmental changes. First, NMDS for each season.

Then PERMANOVA. Let's see PERMANOVA stuff (Table \@ref(tab:PERMANOVAs)).

```{r PERMANOVAs, echo=FALSE}


load('./Outputs/PERMSforALLDF.RData')

emphasize.strong.cells(which(PERMSforAllDF < 0.05, arr.ind = TRUE))
pander(PERMSforAllDF,
             caption = "(\\#tab:PERMANOVAs) The third table.")

# knitr::kable(test,
#              caption = "The first table.")

```

We know the environmental factors are significant, but in what direction do they drive group separation?
RDAs are the answer.
Here they are.

```{r RDAs, echo=FALSE}


load('./Outputs/RDAsigforAll.RData')

# emphasize.strong.cells(which(RD < 0.05, arr.ind = TRUE))
# pander(PERMSforAllDF,
#              caption = "(\\#tab:PERMANOVAs) The third table.")

# knitr::kable(test,
#              caption = "The first table.")

```

You can also embed plots, for example:

```{r deltaTaxa, echo=FALSE, message=FALSE, fig.width=7.5, fig.cap = "Another amazing plot"}

# source("DeltaTaxa.R"
#        , local = knitr::knit_global())
# plot(finalEco)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

### Including Plots {.unnumbered}

### Literature Cited

