---
title: " "
output:
  bookdown::word_document2:
    reference_docx: results_formatting.docx
linkcolor: black
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
library(bookdown)
Sys.setenv(FISHBASE_API="sealifebase")
#knitr::opts_chunk$set(echo = TRUE)
```

### Results  {-}

Richness during summer for all stations was generally stable, while winter was more variable in the southern estuaries than in the northern estuaries (Figure \@ref(fig:richnessSummary)). 

```{r richnessSummary, echo=FALSE, message=FALSE, fig.width=7.5, fig.cap = "An amazing plot"}

source("RichnessSummary.R")
plot(richPlot)

```

However, total abundance over time was quite variable for all systems in the winter season, with summer generally rather stable.

```{r totalAbSummary, echo=FALSE, message=FALSE, fig.width=7.5, fig.cap = "Another amazing plot"}

source("TotalAbundanceSummary.R")
       #, local = knitr::knit_global())
plot(AbSummaryPlot)

```

In [@Miner2021] science happened

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

### Including Plots  {-}

You can also embed plots, for example:

```{r deltaTaxa, echo=FALSE, message=FALSE, fig.width=7.5, fig.cap = "Another amazing plot"}

source("DeltaTaxa.R"
       , local = knitr::knit_global())
plot(finalEco)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### Literature Cited